Notes about Statistics, Clustering, Graphs, \ldots

### 2017-04-19 discussion/discussion.tex
new stuff,experiments,notes

\begin{abstract}
In this document we collect notes about our experiments and theories
so we can discuss together with the group in an organized way.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments using Energy Statistics and One-Dimensional 
Random Projections (Gui 04-01-2017)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Energy Statistics Based Clustering (Gui 04-05-2017)}

@@

>>>>

\begin{abstract}
In this document we collect notes about our experiments and theories
for discussion.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments using Energy Statistics and One-Dimensional 
Random Projections}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Energy Statistics Based Clustering}

@@

\clearpage

\section{1$D$ Energy Clustering}

Now we do energy clustering in $1D$ ($\mathcal{E}1D$), 
which in this case we can optimize
the function exactly. We also compare with the Kernel approach 
(Ker/$\mathcal{E}$)to energy
statistics, and also the graph partitioning approach (EClust). One can see that
these results agree between themselves.

\begin{figure}[h]
\begin{minipage}{.49\textwidth}
\includegraphics[scale=.45]{hist_gauss1.pdf}
\end{minipage}
\begin{minipage}{.49\textwidth}
\begin{tabular}{ l l l l l}
\hline
Ker/$\mathcal{E}$ & $\mathcal{E} 1D$ & EClust & GMM~~ & $k$-means \\
\hline
0.823 &
0.823 &
0.823 &
0.798 &
0.826 \\
0.852 & 
0.852 &
0.852 &
0.852 & 
0.853 \\
0.842 &
0.841 &
0.841 & 
0.842 &
0.839 \\
\hline
\end{tabular}
\end{minipage}
\caption{\label{fig:1dgaus1}
We have $x \sim \tfrac{1}{2}\left( \mathcal{N}(\mu_1, \Sigma_1) +
\mathcal{N}(\mu_2, \Sigma_2)\right)$ where $\mu_1 = 0$, $\mu_2=2$,
$\Sigma_1 = \Sigma_2 = 1$,
and $500$ points on each cluster. We run the experiment three times,
using $k$-means++ as initialization in kernel.
}
\end{figure}

\begin{figure}[h]
\begin{minipage}{.49\textwidth}
\includegraphics[scale=.45]{hist_gauss2.pdf}
\end{minipage}
\begin{minipage}{.49\textwidth}
\begin{tabular}{ l l l l l}
\hline
Ker/$\mathcal{E}$ & $\mathcal{E} 1D$ & EClust & GMM~~ & $k$-means \\
\hline
0.704 &
0.704 &
0.704 &
0.703 &
0.704 \\
0.693 &
0.686 &
0.686 &
0.688 &
0.686 \\
0.687 &
0.688 &
0.688 &
0.691 &
0.691 \\
\hline
\end{tabular}
\end{minipage}
\caption{\label{fig:1dgaus1}
We have $x \sim \tfrac{1}{2}\left( \mathcal{N}(\mu_1, \Sigma_1) +
\mathcal{N}(\mu_2, \Sigma_2)\right)$ where $\mu_1 = 0$, $\mu_2=1$,
$\Sigma_1 = \Sigma_2 = 1$,
and $500$ points on each cluster. We run the experiment three times,
using $k$-means++ as initialization in kernel.
}
\end{figure}

\begin{figure}[h]
\begin{minipage}{.49\textwidth}
\includegraphics[scale=.45]{hist_loggauss1.pdf}
\end{minipage}
\begin{minipage}{.49\textwidth}
\begin{tabular}{ l l l l l}
\hline
Ker/$\mathcal{E}$ & $\mathcal{E} 1D$ & EClust & GMM~~ & $k$-means \\
\hline
0.914 &
0.921 &
0.921 &
0.978 &
0.762 \\
0.894 &
0.894 &
0.894 &
0.979 &
0.8 \\
0.877 &
0.877 &
0.877 &
0.972 &
0.794 \\
\hline
\end{tabular}
\end{minipage}
\caption{\label{fig:1dloggaus1}
We have $x \sim \tfrac{1}{2}\left( \log \mathcal{N}(\mu_1, \Sigma_1) +
\log \mathcal{N}(\mu_2, \Sigma_2)\right)$ where $\mu_1 = 0$, $\mu_2=1.5$,
$\Sigma_1 = 0.25$, $ \Sigma_2 = 0.5$,
and $500$ points on each cluster. We run the experiment three times,
using $k$-means++ as initialization in kernel.
}
\end{figure}

\begin{figure}[h]
\begin{minipage}{.49\textwidth}
\includegraphics[scale=.45]{hist1d_gauss2.pdf}
\end{minipage}
\begin{minipage}{.49\textwidth}
\begin{tabular}{ l l l l l}
\hline
Ker/$\mathcal{E}$ & $\mathcal{E} 1D$ & GMM~~ & $k$-means \\
\hline
0.663&
0.653&
0.606&
0.606\\
0.726&
0.723&
0.793&
0.71 \\
0.663&
0.666&
0.593&
0.66 \\
0.653&
0.62 &
0.653&
0.603\\
0.676&
0.63 &
0.72 & 
0.62 \\
\hline
\end{tabular}
\end{minipage}
\caption{\label{fig:1dloggaus1}
We have $x \sim \tfrac{1}{2}\left( \mathcal{N}(\mu_1, \Sigma_1) +
\mathcal{N}(\mu_2, \Sigma_2)\right)$ where $\mu_1 = 0$, $\mu_2=2$,
$\Sigma_1 = 1$, $ \Sigma_2 = 3$, $n_1 = 100$, $n_2=200$.
For each experiment we run several times and pick the answer with
optimal objective function.
}
\end{figure}

\clearpage

\section{Energy Clustering Higher $D$}

Same as before but in higher dimensions.

\begin{figure}[h]
\begin{minipage}{.49\textwidth}
\includegraphics[scale=.45]{gauss_4d1.pdf}
\end{minipage}
\begin{minipage}{.49\textwidth}
\begin{tabular}{ l l l l}
\hline
Ker/$\mathcal{E}$ & Spectral/$\mathcal{E}$ & GMM~~ & $k$-means \\
\hline
0.7825 &
0.7175 &
0.6 &
0.7675 \\
0.815 &
0.77 &
0.665 &
0.8175 \\
0.8325 &
0.7825 & 
0.685 & 
0.8175 \\
0.835 &
0.7675 &
0.8 & 
0.83 \\
\hline
\end{tabular}
\end{minipage}
\caption{\label{fig:1dloggaus1}
We have $x \sim \tfrac{1}{2}\left( \mathcal{N}(\mu_1, \Sigma_1) +
\mathcal{N}(\mu_2, \Sigma_2)\right)$ where $\mu_1 = (0,0,0,0)^T$, 
$\mu_2=(1.5, 0,0,0)^T$,
$\Sigma_1 = 
\left( \begin{smallmatrix} 
0.6 & 0.2 & 0.2 & 0.2 \\ 
0.2 & 0.6 & 0.2 & 0.2 \\ 
0.2 & 0.2 & 0.6 & 0.2 \\ 
0.2 & 0.2 & 0.2 & 0.6 \\ 
\end{smallmatrix} \right)$, 
$\Sigma_2 = 
\left( \begin{smallmatrix} 
0.5 & 0 & 0 & 0 \\ 
0 & 0.5 & 0 & 0 \\ 
0 & 0 & 0.5 & 0 \\ 
0 & 0 & 0 & 0.5 \\ 
\end{smallmatrix} \right)$,
$n_1=200$, $n_2=200$.
We run the experiments few times and pick the answer with best
objective function.
}
\end{figure}

\begin{figure}[h]
\begin{minipage}{.49\textwidth}
\includegraphics[scale=.45]{loggauss1.pdf}
\end{minipage}
\begin{minipage}{.49\textwidth}
\begin{tabular}{ l l l l}
\hline
Ker/$\mathcal{E}$ & EClust & GMM~~ & $k$-means \\
\hline
0.755 &
0.755 &
0.71 &
0.585 \\
0.74 &
0.745 &
0.81 &
0.575 \\
0.775 &
0.775 &
0.805 &
0.635 \\
\hline
\end{tabular}
\end{minipage}
\caption{\label{fig:1dloggaus1}
We have $x \sim \tfrac{1}{2}\left( \log \mathcal{N}(\mu_1, \Sigma_1) +
\log \mathcal{N}(\mu_2, \Sigma_2)\right)$ where $\mu_1 = (0,0)^T$, 
$\mu_2=(1.5, 0)^T$,
$\Sigma_1 = 
\left( \begin{smallmatrix} 
0.25 & 0 \\ 0 & 0.25 
\end{smallmatrix} \right)$, 
$\Sigma_2 = 
\left( \begin{smallmatrix} 
0.5 & 0 \\ 0 & 0.5 
\end{smallmatrix} \right)$, 
and $100$ points on each cluster. We run the experiment three times,
using $k$-means++ as initialization in kernel. EClust is extremelly slow.
}
\end{figure}

\begin{figure}[h]
\begin{minipage}{.49\textwidth}
\includegraphics[scale=.45]{loggauss2.pdf}
\end{minipage}
\begin{minipage}{.49\textwidth}
\begin{tabular}{ l l l l}
\hline
Ker/$\mathcal{E}$ & EClust & GMM~~ & $k$-means \\
\hline
0.895 &
0.895 &
0.95 &
0.715 \\
0.91 &
0.91 &
0.955 &
0.8 \\
0.815 &
0.815 &
0.93 &
0.69 \\
\hline
\end{tabular}
\end{minipage}
\caption{\label{fig:1dloggaus1}
We have $x \sim \tfrac{1}{2}\left( \log \mathcal{N}(\mu_1, \Sigma_1) +
\log \mathcal{N}(\mu_2, \Sigma_2)\right)$ where $\mu_1 = (0,0,0,0)^T$, 
$\mu_2=(1.5, 1.5,0,0)^T$,
$\Sigma_1 = 
\left( \begin{smallmatrix} 
0.25 & 0 & 0 & 0 \\ 
0 & 0.25 & 0 & 0 \\ 
0 & 0 & 0.25 & 0 \\ 
0 & 0 & 0 & 0.25 \\ 
\end{smallmatrix} \right)$, 
$\Sigma_2 = 
\left( \begin{smallmatrix} 
0.5 & 0 & 0 & 0 \\ 
0 & 0.5 & 0 & 0 \\ 
0 & 0 & 0.5 & 0 \\ 
0 & 0 & 0 & 0.5 \\ 
\end{smallmatrix} \right)$, 
and $100$ points on each cluster. We run the experiment three times,
using $k$-means++ as initialization in kernel. EClust is extremelly slow.
}
\end{figure}

\begin{figure}[h]
\begin{minipage}{.49\textwidth}
\includegraphics[scale=.45]{circles.pdf}
\end{minipage}
\begin{minipage}{.49\textwidth}
\begin{tabular}{ l l l l}
\hline
Ker/$\mathcal{E}$ & EClust & GMM~~ & $k$-means \\
\hline
0.74 &
0.516 &
0.516 &
0.536 \\
0.583 &
0.513 &
0.52 & 
0.556 \\
0.706 &
0.503 &
0.516 &
0.503 \\
0.633 &
0.5 &
0.553 &
0.52 \\
\hline
\end{tabular}
\end{minipage}
\caption{\label{fig:circles}
We generate two circles with random noise.
First circle has radius $1$ and a gaussian noise of size $0.2$.
Second circle has radius $.2$ and a gaussian noise of size $0.2$.
We have 200 points on each class.
}
\end{figure}
